---
title: "Report 1: Survey on students & professors"
output:
  html_document:
    df_print: paged
---

```{r message=FALSE, warning=FALSE}
#Libraries required
library(readr)
library(tidyverse) #includes dplyr

#import the cvs file
professors <- read_csv("M:/proyectos/curso R/Documentacion/zz PROYECTOS/Project 1/professors.csv") %>%
  rename(Questions = `X1`)

students <- read_csv("M:/proyectos/curso R/Documentacion/zz PROYECTOS/Project 1/students.csv") %>%
  rename (Questions = `X1`)

```





#### PART 1 
#### information regarding the number of answers obtained


```{r}
Stotal_answers <- mutate(students,Total_answers_per_question = Disagree + Agree + `Totally agree` + `Totally Disagree`) %>%
  summarise(Stotal_answers = sum(Total_answers_per_question)) %>%
  pull()

Sagree <- students %>%
  summarise(Sagree = sum(Agree)) %>%
  pull()

Sdisagree <- students %>%
  summarise(Sdisagree = sum(Disagree)) %>%
  pull()

Stagree <- students %>%
  summarise(Stagree = sum(`Totally agree`)) %>%
  pull()

Stdisagree <- students %>%
  summarise(Stdisagree = sum(`Totally Disagree`)) %>%
  pull()

```

> **From the students**, we got `r Sagree` agree, `r Sdisagree` disagree, `r Stagree` totally agree, and `r Stdisagree` totally disagree.
This makes a **total amount of `r Stotal_answers` answers** provided.







```{r}
Ptotal_answers <- mutate(professors,Total_answers_per_question = Disagree + Agree + `Totally agree` + `Totally Disagree`) %>%
  summarise(Ptotal_answers = sum(Total_answers_per_question)) %>%
  pull()

Pagree <- professors %>%
  summarise(Pagree = sum(Agree)) %>%
  pull()

Pdisagree <- professors %>%
  summarise(Pdisagree = sum(Disagree)) %>%
  pull()

Ptagree <- professors %>%
  summarise(Ptagree = sum(`Totally agree`)) %>%
  pull()

Ptdisagree <- professors %>%
  summarise(Ptdisagree = sum(`Totally Disagree`)) %>%
  pull()

```

> **From the professors**, we got `r Pagree` agree, `r Pdisagree` disagree, `r Ptagree` totally agree, and `r Ptdisagree` totally disagree.
This makes a **total amount of `r Ptotal_answers` answers** provided.









```{r}

Panswers <- if(mutate(professors,Total_answers_per_question = Disagree + Agree + `Totally agree` + `Totally Disagree`) %>%
   summarise(xDifferent = n_distinct(Total_answers_per_question))  == 1){
  str_glue("all ")
} else {
  str_glue("not all ")
}


```

> On the report, **`r Panswers` questions** have been replied with the same amount of answers.







#### PART 2 
#### side by side charts comparing the distribution of answers, students vs professors

```{r}

joined_tables <- bind_cols(
  (professors %>%
     rename(PA = Agree,
            PD = Disagree,
            PTD = `Totally Disagree`,
            PTA = `Totally agree`)),
  (select(students, -Questions) %>%  #we have already the same Questions index from the Professors table
     rename(SA = Agree,
            SD = Disagree,
            STD = `Totally Disagree`,
            STA = `Totally agree`))
) 



joined_tables %>%
  gather(PA, PD, PTA, PTD, SA, SD, STD, STA, key = "Type_of_response", value = "amount_of_responses")%>%
  ggplot(mapping = aes(x= `Type_of_response`,
                     y=`amount_of_responses`,
                     fill = factor(Type_of_response) )) +
  geom_bar(stat = "identity") +
  facet_wrap("Questions") +                    #groups the outputs based on the "Questions" value
  theme(legend.title = element_blank(),        #removes legend title
        axis.text.x = element_blank(),         #removes label info on axis x
        axis.text.y.left = element_text(size = 7)) +    #changes the text size on the labels on axis y
  
  
  ylab("amount of responses") + xlab("type of responses")+  #axel titles
  ggtitle(">>> distribution of answers <<<") %>%  #main title
  
  view()

```






#### PART 3 
#### MODE for students & professors

```{r}
gathered.data <- joined_tables %>%
  gather(PA, PD, PTA, PTD, SA, SD, STD, STA, key = "Type_of_response", value = "amount_of_responses")

gather.students <- students %>%
  gather(-Questions, key = "Response", value = "frequency") %>%
  arrange(Questions) %>%
  mutate(resp.group = "Student") 

gs <- gather.students %>%
  group_by(Questions) %>% 
  mutate (max.freq = max(frequency)) %>%
  ungroup() %>%
  filter(max.freq == frequency) %>%
  select(-frequency, -resp.group, Mode = Response)
  
  

gather.professors <- professors %>%
  gather(-Questions, key = "Response", value = "frequency") %>%
  arrange(Questions) %>%
  mutate(resp.group = "Student") 

gp <- gather.professors %>%
  group_by(Questions) %>% 
  mutate (max.freq = max(frequency)) %>%
  ungroup() %>%
  filter(max.freq == frequency) %>%
  select(-frequency, -resp.group, Mode = Response)



knitr::kable(gs, caption = "Mode for student's replies")
knitr::kable(gp, caption = "Mode for professor's replies")
```


